```{r,eval=knitr::is_latex_output(),echo=FALSE,results='asis'}
cat("\\FloatBarrier\\newpage")
```

## Appendix {-}

### Appendix A {-}

#### Summary of Data Strategy Implementation (June 2020) {-}

Below are the guidelines used to conduct implementation of the Data Strategy, as of June 2020. This summary describes the Data Strategy Implementation document which is not publicly available.

#### Data Governance {-}

- Making data searchable and managing data quality improves its usability for evidence-informed decision making.
- CCT undertook exercises to understand all data sets and metadata held by the City.
- CCT is developing a data taxonomy of the hierarchical relationships between data sets.
- CCT is developing a data dictionary of terminology and metrics. The data dictionary will standardize usage across units within CCT and improve efficiency in communication and collaboration on data projects.
- An institutionalized data governance structure ensures quality data. Empowering data collectors and generators to take ownership of the data and its quality can improve data cleaning. Human efforts can be augmented by advanced analytics that can detect anomalies and surface uncleaned data.
- CCT also provides standards for data analytics to ensure that the results provided by data analysts can be reproduced and verified.
- CCT uses open-source software, where possible, to scale analytical efforts in collaboration with service providers and academics, to reduce the time constraints of the procurement process, and to reduce the time from design to institutionalization, implementation, and outcome.

#### Data Culture {-}

- CCT's initiatives build institutional knowledge to flexibly manage data and adjust to changes both internally and externally.
- A change management plan is being rolled out to create awareness of, and desire for, working with data.

#### Data Capabilities {-}

- CCT invests in human resources to develop and manage open data architectures.
- Roll-out of training programs will build capacity with data analysts, data scientists, data engineers, data governance experts, data stewards, data custodians, user interface designers, web developers, data collectors, and other staff.
- Introducing a data literacy program will educate the organization as a whole on understanding and communicating with data.
- Data engineering efforts will build the necessary data infrastructure that will bring CCT into the big-data world, with a strong technical foundation for future analysis for decision-making.
- Guidelines on appropriate methods for data gathering and acquisition will be established.
- Strategizing survey efforts will maximize value from data collection, for example by combining household and health surveys into a single exercise.
- Open-source software can increase transparency and decrease the barrier to entry for new staff who often face a steep learning curve to understand proprietary and customized software.
- Sandbox data science and analytics environments are provided to staff, allowing allows them to fail, experiment, and learn.
- The City of Cape Town has expertise in urban monitoring and policy, but lacks the data engineering capacity to transform a data source into a useful tool for managers and policy makers. The City needs to build capacity within existing staff and employ data professionals to augment internal skills and increase data utilization. Cape Town's proximity to three major universities---the University of Cape Town, Stellenbosch University, and the University of the Western Cape---brings hundreds of skilled professionals to CCT.

### Appendix B {-}

#### Research Branch project review criteria based on Research Framework prioritization (June 2020) {-}

This outline of CCT's Research Branch project review criteria summarizes the document, which is not publicly available. The outline is a resource for CCT's strategy and priorities driving research to inform policy.

CCT research focus areas and questions will be prioritized according to certain criteria, including the following:

#### Research that is linked to the core functions of CCT's Constitutional Mandate {-}

- The foundation of the City's research focus is to inform the City's strategy and policy review cycle(s) and sectoral and departmental strategies and plans.
- Local government has been entrusted with public funds and resources in order to fulfil a specific mandate, as envisioned by South Africa's democratic founders and documented in the Constitution. Research questions within this mandate are prioritized.
- It is CCT's responsibility to use its public funds and resources as efficiently and effectively as possible in fulfilling this mandate. Hence, the research in which CCT invests must further this cause.
- CCT's research interests must not be restricted only to what functions it currently undertakes, given that the context in which local government must fulfil its mandate is ever-changing. Research questions can shape CCT's social, economic, and environmental future, so research that is not directly tied to current policy decisions must also be prioritized.
- Research plays an important role in ensuring CCT adapts to new trends and environments by informing future decisions and proposals for action.
- Research topics likely to be of interest to external researchers will be considered. Preference will be given to high-quality doctoral and postdoctoral research.

#### Research that will be utilized to inform current and future decisions {-}

Policy decisions are a function of the decision-making environment and the potential for impact on the wellbeing of Cape Town residents.

##### Decision-making environment {-}

- Research that will inform clear options for decision-makers, providing understanding of complex issues and predicting potential impact of policy decisions, will be prioritized.
- Research will be a decision-making tool when policy options are unclear and both intended and unintended consequences difficult to predict.
- Research can assist to identify and monitor new trends and possible risks.
- Research on key themes can increase CCT's understanding of underlying dynamics on a local issue.
- CCT will prioritize cutting-edge research on high-value topics of interest.

##### The potential impact of the decision{-}

- Research linked to decisions with a greater potential impact will be prioritized.
- Impact can be understood as a function of the following elements:
    a. Budget/Resource implications in both the short- and long-term taking note of the following:
        - Technical debt: Some policies may have lower capital outlays initially but may require technology or a technical approach that will incur significant maintenance costs over time.
        - Opportunity cost: When resources are allocated to implement a policy, what other policies are not implemented? What are the potential benefits for CCT residents of the chosen policy and of the policies that are not chosen?
    b. Risk
        - The extent of known risks associated with a decision and the impact of these risks on the wellbeing of the residents of Cape Town, as well as on the natural environment.
        - The potential for unknown risks associated with a decision, whereby unintended consequences stemming from a decision could yield negative impacts.
    c. Finality
        - The extent to which a decision can be adjusted over time in accordance with feedback received on its impact.
        - Policies and decisions that require high capital investment can be difficult to adjust or course-correct. In such cases, generating information and informing decisions with evidence is key.

#### Research to evaluate the impact of past decisions {-}

- Research that evaluates the impact of a policy decision over time is prioritized according to the budget and resource implications of a policy as well as its associated risks.
- Program and project evaluation are often difficult, particularly when trying to ascertain impact as opposed to outcomes or outputs.
- Funding and resources may be allocated for intensive impact assessments, beyond the scope of the routine Performance Management activities of CCT.

#### Research that supports the City in experimenting with policy or operational responses to complex problems {-}

- Research enables the City to be agile and innovative in response to increasingly complex problems, and to understand the challenging and rapidly changing context in which the City plans, manages, and operates.
- The City will approve research that supports new policy or operational alternatives to assist the City in making decisions and informing actions, with the goal of responsive and effective service delivery.

### Appendix C {-}

<!-- #### Core Metadata Elements for Municipal Datasets (June 2020) {-} -->

These are the metadata elements that are required by general data users in establishing what data exists, how it can be accessed, and potential uses of the data. These are the minimum requirements for adding a dataset to the municipal data catalog.

```{r cctapptabdata, echo=F,results="asis"}
cctapptabdata<- data.frame(
  element = c("Unique ID",
              "Dataset Name",
              "Dataset Description",
              "Dataset Quality",
              "Update Frequency",
              "Data Access Rights",
              "Restricted Reason",
              "Data Format",
              "Data Steward",
              "DS/TR Branch",
              "DS/TR Department",
              "DS/TR Directorate",
              "Data Contact",
              "Data Custodian",
              "Host System ID",
              "Spatial Coverage",
              "Temporal Coverage"),
  description = c("Unique ID for dataset",
                  "A name given to the dataset",
                  "A summary of the dataset, including contents and usage by CCT",
                  "Subjective assessment of dataset quality",
                  "Frequency with which the data are updated in the dataset. Helps researcher understand how often the content is likely to be changed and whether it might be dependent upon a regular event, such as an annual budgetary process",
                  "Classification of the dataset in terms of access",
                  "To be completed if the dataset is classified as Restricted; Indicates the reason for restricting access",
                  "Provides an indication of the interoperable structure of the data as reflected by the primary file format in which the system exports data",
                  "Indicates the individual who assumes business accountability for the dataset in relation to aspects such as access, quality, integrity, and completeness",
                  "Indicates the organizational unit (branch) of the data steward",
                  "Indicates the organizational unit (department) of the data steward",
                  "Indicates the organizational unit (directorate) of the data steward",
                  "Indicates the official who works most closely with, and can best provide additional information about, the data and metadata",
                  "Indicates the official who is responsible for the technical management of the data (e.g., the hosting and serving of the data)",
                  "Unique ID of the application in which the data are stored (primary/source system)",
                  "Indicates areas for which data are recorded",
                  "Indicates the range of time for which data are recorded"),
  notes = c("",
            "Should be short but adequately reflect to what the content of the dataset relates",
            "Should include purpose/usage of the dataset and how the data is sourced/acquired",
            "Should indicate specific gaps or concerns",
            "Historical (not updated), Event Based",
            "Open Public, Internal Open, Internal Restricted, Secret",
            "Personal Information, Security",
            "Examples include: CSV, Relational DB, Excel, Word, PDF, Text, MP4, MPEG",
            "Should be a person who has sufficient authority to authorize decisions in relation to the management of, and access to, the dataset; if the dataset is sourced from an external organization, the data steward should be from the department primarily responsible for obtaining the data on behalf of CCT",
            "",
            "",
            "",
            "",
            "",
            "",
            "Entire city or subareas; Should also indicate the lowest spatial unit at which the data can be presented/analyzed",
            "Specific dates or intervals when data were recorded; Earliest and latest dates for which data were recorded"))

  #   pander::pander(cctapptabdata,
  #                split.cells = c(15,25,25),
  #                justify = 'left',
  #                col.names=c("Element Name","Description","Notes"))

if ( knitr::is_html_output() ) {
   kable(cctapptabdata,booktabs=T,caption="Core Metadata Elements for Municipal Datasets (June 2020)",col.names=c("Element Name","Description","Notes")) %>%
      column_spec(1,width="12em")
}
if ( knitr::is_latex_output() ) {
  cat("\\input{./assets/cct/cctappendixtable.tex}")
}
```


